Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2018-02-02 aggregation/agg-basic-lca.tex

We begin by providing a mathematical formulation of the LCA problem.  An LCA computation is performed over a collection of linked unit processes, each of which describes a single operation or reference product.  The task of procedure can be described in two steps: making an inventory of environmental interventions; and determining their environmental impacts.  If the processes are assumed to be linear and time-invariant, the problem can be written in terms of matrix algebra.  There are two common formulations: the ``technology matrix'' approach and the ``input-output'' or Leontief approach \citep{Heijungs2002, Pauliuk_2015_framework}.  Although the two models are understood to be equivalent, the input-output approach is computationally more efficient in some applications because it lends itself to iterative solution \citep{Peters_JLCA_2007}.  That formulation may be stated:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the \emph{direct requirements matrix}, where each column of values reports the necessary inputs (positive values) and generated non-reference outputs (negative values)\footnote{Whether negative values are present or allowed depends on the particular construct or co-production method being used. \cite{Majeau_Bettez_2014}}
per unit of a process's reference output; $B$ is the environmental intervention or emission matrix; and $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions with respect to a given impact assessment method.  In this model, $\mathbf{y}$ is the externally specified final demand, which is the product or service whose delivery is to be assessed, and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.

The term $(I-A)$ is equivalent to the normalized technology matrix, i.e. the difference between the supply and use tables in classical input-output analysis, $(V'-U)$ \citep{Suh_JIE_2010}.  The conditions under which this matrix has an inverse, particularly with respect to different strategies for handling co-production, have been discussed extensively \citep{Majeau_Bettez_2014}.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}. 
\input{fig_lca}

In classical LCA, the matrices $A$ (or $I-A$) and $B$ are called a life cycle inventory database (LCIDB), and the functional unit of a study is given by the final demand vector $\mathbf{y}$.  Intermediate results of this computation include:
\begin{itemize}
\item $\mathbf{x} = \left(I-A\right)^{-1}\cdot\mathbf{y}$, the activity level vector;
\item $\mathbf{b} = B\cdot\mathbf{x}$, the life cycle inventory;
\item $s = \mathbf{e}^T\cdot\mathbf{b}$, the life cycle impact category indicator, the study result.
\end{itemize}

>>>>

In the most general terms, the LCA problem concerns the interaction of a collection of activities and commodities described by a supply-and-use inventory, in which $U$ is a table of the use of each commodity by each activity (inputs) and $V$ is a table of the supply of each commodity by each activity (outputs).  Such an account can always be transformed into a \textit{symmetric} table of inputs and outputs on a commodity-by-commodity or industry-by-industry basis \citep{Eurostat_2008}. Doing so requires the application of various assumptions and transformations to the data, the details of which are not in scope of this paper, but which have recently been reviewed in depth \citep{Suh_JIE_2010, Majeau_Bettez_2014}.

A symmetric input output table can be written as a table of coefficients, in which each column of the matrix provides a ``recipe'' for creating the corresponding commodity by consuming other commodities (and, if applicable, producing other co-products).  The table of coefficients, commonly known as a \textit{direct requirements matrix}, can be used to compute the output from all activities that is necessary to produce a given commodity.  That vector of outputs can then be combined with a second table of coefficients, known as an \textit{emission matrix}, that reports exchanges with the environment associated with the production of each commodity, to determine the cumulative environmental emissions associated with the production of the commodity over its life cycle.  Those emissions are individually \textit{characterized} with respect to an environmental impact category of interest, and the characterized emissions summed to report the \textit{impact score} or category indicator, which is the result of the study.  The formulation may be summarized:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the direct requirements; $B$ is the emission matrix;  $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions; $\mathbf{y}$ is the externally specified final demand; and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}.
\input{fig_lca}

The term $(I-A)$ is equivalent to a normalized \textit{technology matrix}, and in certain cases can be computed from the difference between the supply and use tables, $(V'-U)$, with equivalent results \citep{Suh_JIE_2010}. 
The coefficient matrices $A$ (or $I-A$) and $B$ together are called a \textit{life cycle inventory database} (LCIDB), and the functional unit of a study is given by the final demand vector $\mathbf{y}$, which is the product or service whose delivery is to be assessed.\footnote{%
Differences in convention have led to both the direct requirements matrix $A$, and the technology matrix, analogous to $(I-A)$, being denoted by the same symbol in different contexts.  If the data contained in the LCIDB is sufficiently rich, either approach may be implemented with equivalent computational results.  The direct requirement formulation has the advantage of permitting iterative solution, which can dramatically reduce the computational resources required \citep{Peters_JLCA_2007}.}  Intermediate results of this computation include:
\begin{itemize}
\item $\mathbf{x} = \left(I-A\right)^{-1}\cdot\mathbf{y}$, the activity level vector;
\item $\mathbf{b} = B\cdot\mathbf{x}$, the life cycle inventory;
\item $s = \mathbf{e}^T\cdot\mathbf{b}$, the life cycle impact category indicator, the study result.
\end{itemize}
